Capstone Movielens Report

### 2025-05-31 reports/capstone-movielens.site/02.8-0.UMGYDE-model.Rmd
Model regularization: support functions described.

###### ***degree***
\
A parameter to pass to the internally called [calc_UMGY_SmoothedDay_effect](#func.calc_UMGY_SmoothedDay_effect) function (see the [Details](#func.train_UMGY_SmoothedDay_effect.details) section below), meaning the degree of the polynomials to be used.

###### ***span***
\
A parameter to pass to the internally called [calc_UMGY_SmoothedDay_effect](#func.calc_UMGY_SmoothedDay_effect) function (see the [Details](#func.train_UMGY_SmoothedDay_effect.details) section below) to control the degree of smoothing.

@@

::: {.noteblock data-latex=""}
The complete source code of building and training the current model is available in the [UMGYDE Model Building](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1432) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1432) script on _GitHub_.
:::

@@

@@

##### [regularize.train_UMGYD_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L97) Function {#func.regularize.train_UMGYD_effect}
\

This is a helper function that calculates the _Smoothed Day Effect_ of the user ratings and is intended to be used in other functions and scripts (such as [regularize.train_UMGYD_effect](#func.regularize.train_UMGYD_effect)) described below.

@@

::: { #func..params. .sidebar }
  - ***train_set:***
The _Train Set_ dataset.

@@

::: { #func..params. .sidebar }
  - ***lambda:***
As explained [above](#sec.umgydem.support_functions.note.lambda) in the [] section, this is the $\lambda$ coefficient of the penalty of the _Regularization_ method.

@@

::: { #func.regularize.train_UMGYD_effect.details .sidebar }
###### Details
\
Under the hood, the function internally calls the 
:::

@@

::: {.noteblock data-latex=""}
The source code of the [regularize.train_UMGYD_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L97) function is also available in the [UMGYDE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L1) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L97) script on _GitHub_.
:::

###### Return
\
A data frame object representing the _User+Movie+Genre+Year+(Smoothed)Day (UMGYD) Effect_.

@@

##### [regularize.train_UMGYD_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L190) Function {#func.regularize.train_UMGYD_effect.cv}
\
The overloaded version of the [train_UMGY_SmoothedDay_effect.RMSE.cv](#func.train_UMGY_SmoothedDay_effect.RMSE.cv) function described above with the fixed [degree](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.params.degree) parameter value of 0.

@@

::: { #func.regularize.train_UMGYD_effect.cv.params.lambda .sidebar }
  - ***lambda:***
As explained [above](#sec.umgydem.support_functions.note.lambda) in the [] section, this is the $\lambda$ coefficient of the penalty of the _Regularization_ method.

@@

::: { #func.regularize.train_UMGYD_effect.cv.details .sidebar }
###### Details
\
Under the hood, the function internally calls the [train_UMGY_SmoothedDay_effect.RMSE.cv](#func.train_UMGY_SmoothedDay_effect.RMSE.cv) described above with the fixed [degree](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.params.degree) parameter value of 0 and the [span](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.params.span) parameter value taken from the [parameter of the same name](#func.regularize.train_UMGYD_effect.cv.params.span) of this function to compute the `RMSE` of the model to be tuned.
:::

###### Source Code
\
Below is the function's source code:
```{r eval=FALSE}
regularize.train_UMGYD_effect.cv <- function(lambda) {
  best_degree <- lss.UMGYDE.best_params["degree"]
  best_span <- lss.UMGYDE.best_params["span"]

@@

::: {.noteblock data-latex=""}
The source code of the [regularize.train_UMGYD_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L190) function is also available in the [Tuning `loess` Parameters](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L170) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L190) script on _GitHub_.
:::

###### Return
\
_Root Mean Square Error (RMSE)_ of the _UMGYD_ model trained using the [stats::loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess) function called with the given [span](#func.regularize.train_UMGYD_effect.cv.params.span) parameter and the `degree` parameter value equal to 0.

@@

::: {.noteblock data-latex=""}
The [regularize.test_lambda.UMGYD_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L240) function are defined in the [Regularization](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L222) section of the [UM-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L240) script on _GitHub.
:::

This function calculates _RMSE_ of the _UMGYDE Model_ using _5-Fold Cross Validation_ for the given $\lambda$ parameter value:

```{r, eval=FALSE}
regularize.test_lambda.UMGYD_effect.cv <- function(lambda){
  if (is.na(lambda)) {
    stop("Function: regularize.test_lambda.UMGY_effect.cv
`lambda` is `NA`")
  }

@@

::: {.noteblock data-latex=""}
Note that we reuse the function [train_UMGYD_effect.cv](#func.train_UMGYD_effect.cv) calling it from the [regularize.test_lambda.UMGYD_effect.cv](#func.regularize.test_lambda.UMGYD_effect.cv), but now with the $\lambda$ parameter different from the default ('lambda = 0') value.
:::

>>>>

::: { #func.train_UMGY_SmoothedDay_effect.params.degree .sidebar }
###### ***degree***
\
A parameter to pass to the internally called [calc_UMGY_SmoothedDay_effect](#func.calc_UMGY_SmoothedDay_effect) function (see the [Details](#func.train_UMGY_SmoothedDay_effect.details) section below), meaning the degree of the polynomials to be used.
:::

::: { #func.train_UMGY_SmoothedDay_effect.params.span .sidebar }
###### ***span***
\
A parameter to pass to the internally called [calc_UMGY_SmoothedDay_effect](#func.calc_UMGY_SmoothedDay_effect) function (see the [Details](#func.train_UMGY_SmoothedDay_effect.details) section below) to control the degree of smoothing.
:::

@@

::: {.noteblock data-latex=""}
The complete source code of building and training the current model is available in the [UMGYDE Model (UMGYDEM) Building](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1745) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1745) script on _GitHub_.
:::

@@

#### UMGYDE Model Regularization: Mathematical Description
\

@@

##### [regularize.train_UMGYD_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L223) Function {#func.regularize.train_UMGYD_effect}
\

The function is an overloaded version of the [train_UMGY_SmoothedDay_effect](#func.train_UMGY_SmoothedDay_effect) function described above, with the fixed values of the [degree](#func.train_UMGY_SmoothedDay_effect.params.degree) and [span](#func.train_UMGY_SmoothedDay_effect.params.span) parameters set to the best values figured out during the tuning process described in the [UMGYDE Model Tuning] section above.

@@

::: { #func.regularize.train_UMGYD_effect.params.train_set .sidebar }
  - ***train_set:***
The _Train Set_ dataset .

@@

::: { #func.regularize.train_UMGYD_effect.params.lambda .sidebar }
  - ***lambda:***
  In accordance with the description in the [UMGYDE Model Regularization: Mathematical Description] section above, this is the _Regularization parameter_ $\lambda$ used for the _UMGYDE Model Regularization_ process.

@@

@@

::: {.noteblock data-latex=""}
The source code of the [regularize.train_UMGYD_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L223) function is also available in the [Regularization](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L222) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L223) script on _GitHub_.
:::

###### Return
\
A data frame object representing the _UMGYD Effect_.

@@

##### [regularize.train_UMGYD_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L232) Function {#func.regularize.train_UMGYD_effect.cv}
\

::: {.noteblock data-latex=""}
The source code of the [regularize.train_UMGYD_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L232) function is defined in the [Regularization](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L222) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L232) script on _GitHub_.
:::

The function is an overloaded version of the [train_UMGY_SmoothedDay_effect.cv](#func.train_UMGY_SmoothedDay_effect.cv) function described above, with the fixed values of the [degree](#func.train_UMGY_SmoothedDay_effect.cv.params.degree) and [span](#func.train_UMGY_SmoothedDay_effect.cv.params.span) parameters set to the best values figured out during the tuning process described in the [UMGYDE Model Tuning] section above.

@@

::: { #func.regularize.train_UMGYD_effect.cv.params.lambda .sidebar }
  - ***lambda:***
  In accordance with the description in the [UMGYDE Model Regularization: Mathematical Description] section above, this is the _Regularization parameter_ $\lambda$ used for the _UMGYDE Model Regularization_ process.

@@

###### Source Code
\
The source code of the function is provided below:
```{r eval=FALSE}
regularize.train_UMGYD_effect.cv <- function(lambda) {
  best_degree <- lss.UMGYDE.best_params["degree"]
  best_span <- lss.UMGYDE.best_params["span"]

@@

::: {.noteblock data-latex=""}
Note that we reuse the function [train_UMGY_SmoothedDay_effect.cv](#func.train_UMGY_SmoothedDay_effect.cv) calling it internally from the [regularize.train_UMGYD_effect.cv](#func.regularize.train_UMGYD_effect.cv), but now with the $\lambda$ parameter different from the default ('lambda = 0') value.
:::

###### Return
\
A data frame object representing the _UMGYD Effect_.

@@

The function calculates _RMSE_ of the _UMGYDE Model_ using _5-Fold Cross Validation_ for the given $\lambda$ parameter value.

::: { #func.regularize.test_lambda.UMGYD_effect.cv.params.lambda .sidebar }
  - ***lambda:***
  In accordance with the description in the [UMGYDE Model Regularization: Mathematical Description] section above, this is the _Regularization parameter_ $\lambda$ used for the _UMGYDE Model Regularization_ process.

::: { #func.regularize.test_lambda.UMGYD_effect.cv.details .sidebar }
##### Details
\
Under the hood, the function internally calls the [regularize.train_UMGYD_effect.cv](#func.regularize.train_UMGYD_effect.cv) and then [calc_UMGY_SmoothedDay_effect.RMSE.cv](#func.calc_UMGY_SmoothedDay_effect.RMSE.cv) described above, and returns the _RMSE_ value that the latter inner function calculates for the given [lambda](#func.regularize.test_lambda.UMGYD_effect.cv.params.lambda) parameter value passed to the former one.
:::

###### Source Code
\
The source code of the function is provided below:
```{r, eval=FALSE}
regularize.test_lambda.UMGYD_effect.cv <- function(lambda){
  if (is.na(lambda)) {
    stop("Function: regularize.test_lambda.UMGY_effect.cv
`lambda` is `NA`")
  }

@@

::: {.noteblock data-latex=""}
The source code of the [regularize.test_lambda.UMGYD_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L240) function is also available in the [Regularization](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L222) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L240) script on _GitHub_.
:::

###### Return
\
_Root Mean Squared Error (RMSE)_ of the _UMGYD Model_ trained for the given [lambda](#func.regularize.train_UMGYD_effect.cv.params.span) parameter value.
